
%% ============================================================
\appendix
\section{Supplementary: Infrastructure and Pipeline}
\label{sec:appendix-pipeline}
%% ============================================================

\newcommand{\cmd}[1]{\colorbox{gray!10}{\texttt{#1}}}
\newcommand{\field}[1]{\texttt{\textbf{#1}}}

%% ============================================================
\subsection{Infrastructure and Dependencies}
%% ============================================================

What tools we use for A3 and why each was chosen.

\begin{itemize}[nosep]
  \item Nanochat fork --- training codebase
  \item Modal --- GPU compute
  \item Weights \& Biases --- experiment tracking
  \item YAML configs --- run specification
\end{itemize}

\subsubsection{Nanochat Fork --- Training Code}

We fork Karpathy's nanochat (Feb 2026) so that P2 can modify the architecture without touching the original code.

\textbf{URL}: \url{https://github.com/seonghyunban/nanochat}

\paragraph{Branches}

\begin{itemize}[nosep]
  \item \texttt{master} --- original nanochat, frozen. All baseline runs use this. Do not push here.
  \item \texttt{p2} --- P2's architecture changes, branched from \texttt{master}.
\end{itemize}

\paragraph{Tag}

\begin{itemize}[nosep]
  \item \texttt{baseline-v0} --- immutable snapshot of \texttt{master}. Safety net: even if someone pushes to \texttt{master}, this tag never moves.
\end{itemize}

\paragraph{How Modal Uses the Fork}

The container clones the fork once at image build and caches it. Before every run, it does \texttt{git fetch} to pull new commits, then checks out the branch/tag the config asks for. No container rebuild needed when P2 pushes new code.

\subsubsection{Modal --- GPU Compute}

A3 needs single-GPU jobs. Modal is the simplest platform for this: write a Python function, pick a GPU, run it. \$500 credits ready via promo code, per-second billing, and Volumes to save checkpoints between runs.

\paragraph{Why Not Others}

Anyscale (\$1{,}000) is built around Ray for distributed workloads --- more setup than needed for single-GPU jobs, kept as P4 backup. AWS (\$500) has the most overhead for simple jobs (EC2, SSH, drivers, storage). Snowflake (\$400 trial) is a data warehouse, not a training platform.

\paragraph{Which GPU}

Nanochat requires BF16, so A10G is the minimum. A100/H100 are faster and often cheaper in total cost due to shorter wall time.

\begin{table}[H]
\centering\small
\begin{tabular}{@{}llrr@{}}
\toprule
Part & GPU & Est.\ Time & Est.\ Cost \\
\midrule
P2 (4 ablation runs) & A100 & 1--5 hrs & \$3--13 \\
P3 (2 training stages + eval) & A100 & 1--3 hrs & \$3--8 \\
P4 (full nanochat) & H100 & 6--20 hrs & \$24--79 \\
\bottomrule
\end{tabular}
\end{table}

\subsubsection{Weights \& Biases --- Experiment Tracking}

Dashboard for training curves, loss values, and run configs. Already built into nanochat --- no extra code needed.

\begin{itemize}[nosep]
  \item \textbf{Project}: \texttt{490-autobook-a3} (shared across all parts)
  \item \textbf{Tags}: each run tagged \texttt{p2}, \texttt{p3}, or \texttt{p4}
  \item \textbf{Git hash}: every run logs which nanochat commit it used
\end{itemize}

\subsubsection{YAML Configs --- Run Specification}

Each experiment is a YAML file whose fields map directly to nanochat's command-line flags (\texttt{--depth}, \texttt{--max-seq-len}, etc.). No extra config framework needed --- nanochat already parses its own arguments.

\begin{itemize}[nosep]
  \item \texttt{nanochat\_ref} field picks which branch or tag to run
  \item Comments in the YAML explain each field --- no separate docs needed
\end{itemize}

%% ============================================================
\subsection{Getting Started}
%% ============================================================

One-time setup, how the pipeline works, and the nanochat fork workflow.

\subsubsection{Prerequisites and One-Time Setup}

Three accounts needed: GitHub (you already have this), Modal (\url{https://modal.com}) for GPU compute, and Weights \& Biases (\url{https://wandb.ai}) for experiment tracking. Create your Modal and W\&B accounts, then send your W\&B username to a teammate so they can add you to project \texttt{490-autobook-a3}. Modal gives \$30 free credits on signup; we also have a shared \$500 academic credit pool.

\paragraph{Install W\&B and get your API key}

Run \cmd{pip install wandb}. Find your API key at \url{https://wandb.ai/authorize} --- you will need it in the next step. You also need the local package to fetch training metrics (loss, bpb) for your writeup.

\paragraph{Install Modal CLI and authenticate}

Run \cmd{pip install modal}, then \cmd{modal token new} (opens browser to authenticate). Create a secret so Modal can log to W\&B on your behalf: \cmd{modal secret create wandb-secret WANDB\_API\_KEY=<your-key>}. Verify with \cmd{modal secret list} --- should show \cmd{wandb-secret}.

\subsubsection{How the Pipeline Works}

You write a YAML config. The pipeline reads it, sends training and eval jobs to Modal GPUs, and writes results to a Modal Volume and W\&B. You never need to edit the pipeline code --- everything is controlled via your config.

\paragraph{Files and call graph}

\begin{forest}
  for tree={
    font=\ttfamily\small,
    grow'=0,
    child anchor=west,
    parent anchor=south,
    anchor=west,
    calign=first,
    edge path={
      \noexpand\path [draw, -{|}] (!u.south west) +(7.5pt,0) |- node[fill,inner sep=1.25pt] {} (.child anchor)\forestoption{edge label};
    },
    before typesetting nodes={
      if n=1 {insert before={[,phantom]}} {}
    },
    fit=band,
    before computing xy={l=15pt},
    s sep=4pt,
  }
  [a3/shared/scripts/
    [main.py \normalfont{\textit{--- parses YAML, calls} \texttt{run.remote()}}]
    [shared/
      [infra.py \normalfont{\textit{--- app, image, volume, secrets, setup()}}]
      [helpers.py \normalfont{\textit{--- pure Python utilities}}]
    ]
    [runners/
      [run.py \normalfont{\textit{--- orchestrator: spawns stages then evals}}]
      [train.py \normalfont{\textit{--- one training stage per GPU}}]
      [evaluate.py \normalfont{\textit{--- one checkpoint eval per GPU}}]
    ]
  ]
\end{forest}

\vspace{6pt}
\begin{center}
\definecolor{powderblush}{HTML}{E8C4C4}
\definecolor{parchment}{HTML}{F1E9D2}
\definecolor{silver}{HTML}{C0C0C0}
\definecolor{greyolive}{HTML}{A2A38B}
\definecolor{taupe}{HTML}{8B7E74}
\definecolor{sunflowergold}{HTML}{FFD700}
\definecolor{darkteal}{HTML}{005F5F}
\begin{tikzpicture}[
  node distance=5mm and 6mm,
  base/.style={draw=taupe, rounded corners=2pt, minimum height=6mm, font=\small\ttfamily},
  pyfile/.style={base, fill=powderblush},
  worker/.style={base, fill=greyolive!30, minimum width=14mm},
  wbnode/.style={base, fill=sunflowergold!50, minimum width=18mm},
  volnode/.style={base, fill=darkteal!25, minimum width=18mm},
  arr/.style={-{Stealth[length=2.5mm]}, thick, color=taupe},
  dataarr/.style={-{Stealth[length=2mm]}, thick, color=taupe, opacity=0.3},
  lbl/.style={font=\sffamily\scriptsize\itshape, text=black!60},
]
  % === LOCAL ===
  \node[pyfile] (yaml) {your.yaml};
  \node[pyfile, right=8mm of yaml] (main) {main.py};
  \draw[arr] (yaml) -- (main);
  \node[lbl, above=1pt of main, text=white] {parses YAML};

  % === MODAL: orchestrator ===
  \node[pyfile, right=30mm of main, minimum width=16mm] (run) {run.py};
  \draw[arr] (main) -- node[lbl, above] {.remote()} (run);
  \node[lbl, above=1pt of run] {orchestrator};

  % W&B
  \node[wbnode, right=60mm of run] (wandb) {W\&B};
  \node[lbl, below=1pt of wandb] {\scriptsize curves};

  % Setup
  \node[worker, below=8mm of run] (setup) {setup};
  \draw[arr] (run) -- (setup);
  \node[lbl, right=2mm of setup] {tokenizer + data};

  % train.py
  \node[pyfile, minimum width=20mm, right=4.5mm of setup, yshift=-12mm] (trainpy) {train.py};

  % evaluate.py
  \node[pyfile, minimum width=20mm, below=22mm of trainpy] (evalpy) {evaluate.py};

  % Spine
  \draw[color=taupe, thick] (setup.south) -- (setup.south |- evalpy.west);
  \draw[arr] (setup.south |- trainpy.west) -- (trainpy.west);
  \draw[arr] (setup.south |- evalpy.west) -- (evalpy.west);

  % Trains
  \node[worker, right=8mm of trainpy, yshift=8mm] (t1) {train 1};
  \node[worker, below=2mm of t1] (t2) {train 2};
  \node[worker, below=2mm of t2] (t3) {train 3};
  \coordinate (tfork) at ([xshift=4mm]trainpy.east);
  \draw[color=taupe, thick] (trainpy.east) -- (tfork);
  \draw[arr] (tfork) |- (t1.west);
  \draw[arr] (tfork) -- (t2.west);
  \draw[arr] (tfork) |- (t3.west);

  % Evals
  \coordinate (eval-anchor) at ([yshift=8mm]evalpy.east);
  \node[worker, anchor=center] at (t1 |- eval-anchor) (e1) {eval 1};
  \node[worker, below=2mm of e1] (e2) {eval 2};
  \node[worker, below=2mm of e2] (e3) {eval 3};
  \coordinate (efork) at ([xshift=4mm]evalpy.east);
  \draw[color=taupe, thick] (evalpy.east) -- (efork);
  \draw[arr] (efork) |- (e1.west);
  \draw[arr] (efork) -- (e2.west);
  \draw[arr] (efork) |- (e3.west);

  % Volume
  \coordinate (mid) at ($(t3.south)!0.5!(e1.north)$);
  \node[volnode] at (wandb |- mid) (vol) {Volume};
  \node[lbl, below=1pt of vol] {\scriptsize checkpoints, CSVs};

  % Data arrows
  \draw[dataarr] (t1.east) to[out=15, in=170] (wandb.west);
  \draw[dataarr] (t2.east) to[out=10, in=180] (wandb.west);
  \draw[dataarr] (t3.east) to[out=5, in=190] (wandb.west);
  \draw[dataarr] (t1.east) to[out=-5, in=150] (vol.west);
  \draw[dataarr] (t2.east) to[out=-10, in=160] (vol.west);
  \draw[dataarr] (t3.east) to[out=-15, in=170] (vol.west);
  \draw[dataarr] (e1.east) to[out=15, in=190] (vol.west);
  \draw[dataarr] (e2.east) to[out=10, in=200] (vol.west);
  \draw[dataarr] (e3.east) to[out=5, in=210] (vol.west);

  % Background boxes
  \begin{scope}[on background layer]
    \definecolor{localbox}{HTML}{1B1B1B}
    \node[draw=localbox, rounded corners=4pt, fill=localbox,
          fit=(yaml)(main),
          inner sep=5mm, label={[font=\scriptsize\color{parchment!40!black}]above:local}] {};
  \end{scope}
  \begin{scope}[on background layer]
    \node[draw=silver!60!black, dashed, rounded corners=4pt, fill=silver!20,
          fit=(run)(setup)(trainpy)(t1)(t3)(evalpy)(e1)(e3)(wandb)(vol),
          inner sep=5mm, label={[xshift=13mm, anchor=south east, font=\scriptsize\color{silver!30!black}]above right:Modal cloud}] {};
  \end{scope}
\end{tikzpicture}
\end{center}

\paragraph{Workflow}

From your perspective, using the pipeline is two steps:

\begin{enumerate}[nosep]
  \item \textbf{Write a YAML config} describing what to train and what to evaluate. The config specifies your model depth, sequence length, GPU type, W\&B run name, and which evals to run.
  \item \textbf{Run one command}, passing your config as a parameter:
\end{enumerate}

\begin{lstlisting}
modal run --detach a3/shared/scripts/main.py --config a3/<part>/configs/<your-config>.yaml
\end{lstlisting}

That's it. Modal spins up GPU containers, trains your model, runs evals, and writes results to a Modal Volume (checkpoints, CSVs, JSONs) and W\&B (training curves). You can close your laptop --- \texttt{--detach} keeps the pipeline running on Modal. Check progress on the Modal dashboard or W\&B.

\subsubsection{Nanochat Fork}

Nanochat is an LLM training codebase by Karpathy. We maintain a fork at \url{https://github.com/seonghyunban/nanochat} so that teammates can modify the architecture without touching the original code. The pipeline clones this fork into every Modal container and checks out whichever branch or tag your config specifies.

\paragraph{How to modify vanilla nanochat}

\begin{enumerate}[nosep]
  \item Clone the fork: \texttt{git clone https://github.com/seonghyunban/nanochat.git}
  \item Create a branch: \texttt{git checkout -b p2}
  \item Make your architecture changes, commit, and push: \texttt{git push origin p2}
\end{enumerate}

Never push to \texttt{master} --- it is frozen as the unmodified baseline.

\paragraph{How to use your modified nanochat}

Set \texttt{nanochat\_ref} in your YAML config to your branch name:

\begin{lstlisting}
nanochat_ref: p2
\end{lstlisting}

\textbf{Push before you run.} Modal runs \texttt{git fetch} + \texttt{checkout} on the remote fork --- it does not see your local changes. Forgetting to push means Modal silently runs old code.

\paragraph{How to use vanilla nanochat}

Use the immutable baseline tag:

\begin{lstlisting}
nanochat_ref: baseline-v0
\end{lstlisting}

This tag points to unmodified nanochat and never moves, even if someone pushes to \texttt{master}.

%% ============================================================
\subsection{Running, Monitoring \& Troubleshooting}
%% ============================================================

How to launch a training run, check its progress, collect results, and fix common issues.

\subsubsection{Running}

\begin{lstlisting}
modal run --detach a3/shared/scripts/main.py --config a3/<part>/configs/<name>.yaml
\end{lstlisting}

The \cmd{-{}-detach} flag sends the entire pipeline to Modal's cloud so it survives laptop close, Wi-Fi drops, and terminal exit. Without it, output streams to your terminal but the run dies if you disconnect. Always use \cmd{-{}-detach} for real runs; omit it only for quick debugging where you want immediate output.

\subsubsection{Monitoring}

\paragraph{On the website}

\textbf{Modal dashboard} (\url{https://modal.com}): go to your workspace $\to$ \textbf{Apps} $\to$ \cmd{a3-nanochat}. You can see running, completed, and failed function calls. Click any function call and open the \textbf{Logs} tab for live output. The pipeline prints \cmd{[pipeline]}, \cmd{[train]}, and \cmd{[eval]} prefixed lines so you can follow progress.

\textbf{W\&B} (\url{https://wandb.ai}): go to project \cmd{490-autobook-a3} and find your run by the \field{run} name from your YAML config. The \textbf{Charts} tab shows \cmd{train/loss} and \cmd{val/bpb} over steps in real time.

\paragraph{On the terminal}

You can monitor everything without leaving the terminal.

\textbf{Check if your run is still going:}

\begin{lstlisting}
modal app list
\end{lstlisting}

\textbf{Stream live logs from a detached run:}

\begin{lstlisting}
modal app logs <app-id>
\end{lstlisting}

Reconnects to a running (or recently finished) app and streams its output. The app ID is printed when you launch with \cmd{-{}-detach}.

\textbf{Stop a run:}

\begin{lstlisting}
modal app stop <app-id>
\end{lstlisting}

\subsubsection{Results}

\paragraph{On the website}

\textbf{W\&B}: training curves (\cmd{train/loss}, \cmd{val/bpb}) are on the \textbf{Charts} tab. The \textbf{Config} tab shows all training args and \cmd{git\_hash} for reproducibility.

\textbf{Modal Volume}: browse at \url{https://modal.com} $\to$ your workspace $\to$ \textbf{Volumes} $\to$ \cmd{a3-checkpoints}.

\paragraph{On the terminal}

All eval outputs are stored on your Modal Volume (\cmd{a3-checkpoints}).

\textbf{Browse files:}

\begin{lstlisting}
modal volume ls a3-checkpoints base_checkpoints/         # checkpoints
modal volume ls a3-checkpoints base_eval/                # CORE CSVs
modal volume ls a3-checkpoints custom_evals/             # custom eval JSONs
\end{lstlisting}

\textbf{Download files:}

\begin{lstlisting}
modal volume get a3-checkpoints base_eval/<file> ./results/
modal volume get a3-checkpoints custom_evals/<file> ./results/
\end{lstlisting}

\textbf{Fetch W\&B metrics} (loss, bpb) via Python:

\begin{lstlisting}[language=Python]
import wandb
api = wandb.Api()
run = api.run("<team>/490-autobook-a3/<run-id>")
df = run.history()       # pandas DataFrame with loss, bpb, etc.
df.to_csv("metrics.csv")
\end{lstlisting}

\subsubsection{Troubleshooting}

\paragraph{OOM (out of memory)}
\textbf{Symptom}: \cmd{CUDA out of memory} in logs.\\
\textbf{Fix}: halve \field{device\_batch\_size} in your YAML (e.g.\ 512 $\to$ 256 $\to$ 128) and rerun.

\paragraph{Secret not found}
\textbf{Symptom}: \cmd{Secret 'wandb-secret' not found}.\\
\textbf{Fix}: run \cmd{modal secret list} and check that the name is exactly \cmd{wandb-secret}. If missing, create it (see Prerequisites above).

\paragraph{Checkpoint not found}
\textbf{Symptom}: \cmd{No checkpoints found in ...}\\
\textbf{Fix}: run \cmd{modal volume ls a3-checkpoints base\_checkpoints/} and verify that \field{model\_tag} in your eval config matches a directory that exists.

\paragraph{Git ref not found}
\textbf{Symptom}: \cmd{error: pathspec '<ref>' did not match ...}\\
\textbf{Fix}: push your branch to the nanochat fork first. Verify it appears on GitHub before rerunning.

\paragraph{Timeout}
\textbf{Symptom}: run killed mid-training, Modal dashboard shows timeout.\\
\textbf{Fix}: increase \field{timeout\_hours} in your YAML config and rerun.
